---
title: "Trend of Health Expenditure by Providers In USA"
author: "Abid Al Reza"
date: "7/30/2020"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)
```

## Data Source

This presentation contains a plot created using the function `plot_ly` which depicts the trend of health care expenditure by different type care providers in USA from the year 1980 to 2014.

Different type care providers include different type of establishment delivering care (hospitals, physicians and clinics, nursing homes, etc.) and medical products (prescription drugs, over-the-counter medicines, and sundries and for durable medical products such as eyeglasses and hearing aids) purchased in retail outlets.


The data was collected from [here](https://www.cms.gov/Research-Statistics-Data-and-Systems/Statistics-Trends-and-Reports/NationalHealthExpendData/NationalHealthAccountsStateHealthAccountsProvider).

```{r data processing}
# library loading
if(!require(dplyr)){install.packages('dplyr')}; library(dplyr)
if(!require(tidyr)){install.packages('tidyr')}; library(tidyr)
if(!require(plotly)){install.packages('plotly')}; library(plotly)

# data download and loading
url <- "https://www.cms.gov/Research-Statistics-Data-and-Systems/Statistics-Trends-and-Reports/NationalHealthExpendData/Downloads/provider-state-estimates.zip"
temp <- tempfile()
download.file(url, temp)
mydata <- read.csv(unz(temp, "PROV_US_AGGREGATE14.CSV"), header = TRUE)
unlink(temp)

# data processing 
mydata1<- mydata %>% 
    filter(Group == "United States") %>%
    filter(Item != "Personal Health Care (Millions of Dollars)")%>%
    select(Item,starts_with("Y"))%>% 
    pivot_longer(cols = starts_with("Y"),
                 names_to= "year",
                 values_to="Millions_of_Dollars")%>%
    mutate(year = as.numeric(gsub("Y","",year)))%>%
    mutate(Item = gsub(" +\\(.*\\)","",Item))
```


```{r plot code}
# plot_ly plot code
health_plot<- plot_ly() %>%
    add_trace(data = mydata1,
              x=~year,
              y=~Millions_of_Dollars,
              color=~Item,
              text = ~Item,
              colors = "Paired",
              type = 'scatter', 
              mode = 'lines+markers',
              hovertemplate = paste(
                  "<b>%{text}</b><br>",
                  "%{y:,.0f} Millions USD<br>",
                  "%{xaxis.title.text}: %{x}",
                  "<extra></extra>"
              ))%>%
    layout(title = "Trend of Health Expenditure by Providers In USA",
           xaxis = list(title="Year"),
           yaxis = list(title="Millions of Dollars"),
           hovermode = "x unified")%>%
    layout(legend = list(orientation = 'h',x=0.1, y = -0.2))
```

## Plot
```{r plot}
health_plot
```

